### 2020-05-24 discussion.tex
mostly typos and two graphs

Reliability-aware resource scheduling also has its limitations, as the network
architecture needs to be taken into account. Titan's 3D torus network creates
certainly more challenges for efficient job allocations, than the fat tree
network of Titan's successor, Summit~\cite{olcf:summit}. Nodes associated with
the same job need to be close to each other on the network for maximum
application performance. Also, node outages or less reliable nodes within a
tight group of reliable nodes create resource allocation
holes. Saparating two lower nodes in each cage 0 on Titan would not
create a contiguous partition, yet this group would be the most
reliable in 2016 as was shown in the preceeding section.
%
Alternatively, partially or completely replacing the aging and failing
supercomputer with a new system, even if this solution offers less capability,
may be more cost effective in the end.

>>>>

Reliability-aware resource scheduling also has its limitations, as the network
architecture needs to be taken into account. Titan's 3D torus network creates
certainly more challenges for efficient job allocations, than the fat tree
network of Titan's successor, Summit~\cite{olcf:summit}. Nodes associated with
the same job need to be close to each other on the network for maximum
application performance. Also, node outages or less reliable nodes within a
tight group of reliable nodes create resource allocation
holes. Separating two lower nodes in each cage 0 on Titan would not
create a contiguous partition, yet this group would be the most
reliable in 2016 as was shown in the preceeding section.
%
Alternatively, partially or completely replacing the aging and failing
supercomputer with a new system, even if this solution offers less capability,
may be more cost effective in the end.

### 2020-05-24 intro.tex
mostly typos and two graphs

We begin by describing related work in Sec.~\ref{section:related}. Our
data collection is desctibed in Sec.~\ref{section:dataprep} and the
data cleaning, checking, and lifetime summarization process in
Sec.~\ref{section:dataclean}. A time between failures (TBF) analysis
is given in Sec.~\ref{section:tbf}. Survival analysis based on
Kaplan-Meier modeling (KM) and Cox regression modeling (CR) of
relative hazard rates in various locations are in
Sec.~\ref{section:survival}. The goals of a TBF analysis and of
survival analysis (SA) are different: TBF studies largely how GPU
reliability affects the machine, wheras SA studies largely how the
machine affects GPU reliability.  Finally, we discuss our main
conclusions and recommendations in Sec.~\ref{section:discussion}.

>>>>

We begin by describing related work in Sec.~\ref{section:related}. Our
data collection is described in Sec.~\ref{section:dataprep} and the
data cleaning, checking, and lifetime summarization process in
Sec.~\ref{section:dataclean}. A time between failures (TBF) analysis
is given in Sec.~\ref{section:tbf}. Survival analysis based on
Kaplan-Meier modeling (KM) and Cox regression modeling (CR) of
relative hazard rates in various locations are in
Sec.~\ref{section:survival}. The goals of a TBF analysis and of
survival analysis (SA) are different: TBF studies largely show GPU
reliability affects the machine, whereas SA studies largely how the
machine affects GPU reliability.  Finally, we discuss our main
conclusions and recommendations in Sec.~\ref{section:discussion}.

### 2020-05-24 related.tex
mostly typos and two graphs

Tiwari et al~\cite{Tiwari15Experience,7056044} analyze types of GPU errors on Titan and on a GPU cluster at Los Alamos National Laboratory. This also includes neutron beam experiments at the Los Alamos Neutron Science Center and at Rutherford Appleton Laboratories to further understand GPU soft errors caused by neutron radiation triggered upsets (bit flips and timing errors).
%
Tiwari et al~\cite{10.1145/2807591.2807666} further analyze types of GPU errors on Titan, including software/firmware related errors and failures, ECC double-bit errors, and GPU ``off the bus'' and ECC page retirement events. This work observes an ECC double-bit mean-time between errors (MTBE) of about 160 hours, or about one per week, with 86\% occurring in GPU memory and the rest in the GPU register file. The GPU ``off the bus'' events were caused by a system integration issue that was fixed and not an inherent GPU or GPU memory architecture flaw.
%
Nie et al.~\cite{7446091,nie17characterizing,nie18machine} characterize the soft error behavior of Titan's GPUs in relation to temperature and power consumption to predict their increased occurrence by correlating data in temporal and spatial domains and using machine learning models. The work primarily focuses on correctable single-bit errors.
%
Tiwari et al. and Nie et al. d0 not address the atypical failure mode discussed in this paper. They also only use Titan error and failure data from June 2013 to February 2015, excluding early failures and the failure mode discussed in this paper (which occurred after February 2015).

@@

Other work in characterizing supercomputer faults, errors and failures focuses on the various other systems deployed around the world, primarily at US Department of Energy laboratories.
%
Di et al.~\cite{8809553} develop an in-depth understanding of failure characteristics of the IBM Blue Gene/Q Mira system at Argonne National Laboratory. This work shows that 99.4\% of job failures are due to user behavior, such as errors in the code or misconfiguration. Martino et al.~\cite{6903615} characterize the errors and failures of the Blue Waters Cray XE6/XK7 system at the National Center for Supercomputing Applications (NCSA), at the University of Illinois at Urbana-Champaign. The results show that 74.4\% of system-wide outages in Blue Waters were caused by software. Blue Waters's XK7 partition had the same architecture as Titan, but did not experience the same GPU failure mode detailed in this paper.
% Interesting ... do we know why?

>>>>

Tiwari et al~\cite{Tiwari15Experience,7056044} analyze types of GPU errors on Titan and on a GPU cluster at Los Alamos National Laboratory. This also includes neutron beam experiments at the Los Alamos Neutron Science Center and at Rutherford Appleton Laboratories to further understand GPU soft errors caused by neutron radiation triggered upsets (bit flips and timing errors).
%
Tiwari et al~\cite{10.1145/2807591.2807666} further analyze types of GPU errors on Titan, including software/firmware related errors and failures, ECC double-bit errors, and GPU ``off the bus'' and ECC page retirement events. This work observes an ECC double-bit mean-time between errors (MTBE) of about 160 hours, or about one per week, with 86\% occurring in GPU memory and the rest in the GPU register file. The GPU ``off the bus'' events were caused by a system integration issue that was fixed and not an inherent GPU or GPU memory architecture flaw.
%
Nie et al.~\cite{7446091,nie17characterizing,nie18machine} characterize the soft error behavior of Titan's GPUs in relation to temperature and power consumption to predict their increased occurrence by correlating data in temporal and spatial domains and using machine learning models. The work primarily focuses on correctable single-bit errors.
%
Tiwari et al. and Nie et al. do not address the atypical failure mode discussed in this paper. They also only use Titan error and failure data from June 2013 to February 2015, excluding early failures and the failure mode discussed in this paper (which occurred after February 2015).

@@

Other work in characterizing supercomputer faults, errors and failures focuses on the various other systems deployed around the world, primarily at US Department of Energy laboratories.
%
Di et al.~\cite{8809553} develop an in-depth understanding of failure characteristics of the IBM Blue Gene/Q Mira system at Argonne National Laboratory. This work shows that 99.4\% of job failures are due to user behavior, such as errors in the code or misconfiguration. Martino et al.~\cite{6903615} characterize the errors and failures of the Blue Waters Cray XE6/XK7 system at the National Center for Supercomputing Applications, at the University of Illinois at Urbana-Champaign. The results show that 74.4\% of system-wide outages in Blue Waters were caused by software. Blue Waters's XK7 partition had the same architecture as Titan, but did not experience the same GPU failure mode detailed in this paper.
% Interesting ... do we know why?
